
## Data preparation

Сonvert data to csv format to read it easily with R and remove unuseful columns from it

```{r}
data <- read.csv("../data/IAD30284_384WellPlateDataSheet.csv", header = TRUE, skip = 14)
data <- data[data$Amplicon_ID != "Blank", ]
data$X384Well_Row <- NULL
data$X384Well_Col <- NULL
data$Gene_Symbol <- NULL
data$Genome_Version <- NULL
data$LineItem_Type <- NULL
```

Get amplicon sequence that lies between primers
```{r}
library(jsonlite)
library(httr)
n <- nrow(data)
chromosomes <- data$Chr
beginings <- data$Insert_Start
endings <- data$Insert_Stop
ampliconInnerPart <- rep("", n)
for (i in 1:n){
  call <- paste(c("http://178.79.134.178/refservice/references/GRCh37.p13/sequence?contig=", as.character(chromosomes[i]),
                         "&start=",as.character(beginings[i]),"&end=",as.character(endings[i])), collapse = "")
  call
  jsonData <- fromJSON(call)
  ampliconInnerPart[i] <- jsonData$sequence
}
head(ampliconInnerPart)
data$Amplicon_Inner_Part <- ampliconInnerPart
```

Get dangling ends
```{r}
danglingEndStart <- rep("", n)
danglingEndStop <- rep("", n)
for (i in 1:n){
  call <- paste(c("http://178.79.134.178/refservice/references/GRCh37.p13/sequence?contig=", as.character(chromosomes[i]),
                         "&start=",as.character(beginings[i]-1),"&end=",as.character(beginings[i]-1)), collapse = "")
  jsonData <- fromJSON(call)
  danglingEndStart[i] <- jsonData$sequence
  call <- paste(c("http://178.79.134.178/refservice/references/GRCh37.p13/sequence?contig=", as.character(chromosomes[i]),
                         "&start=",as.character(endings[i]+1),"&end=",as.character(endings[i]+1)), collapse = "")
  jsonData <- fromJSON(call)
  danglingEndStop[i] <- jsonData$sequence
}
data$Start_Dangling_End <- danglingEndStart
data$Stop_Dangling_End <- danglingEndStop
```

Load results and get a proportion for each
```{r}
result <- read.csv("../data/result.csv", header = TRUE)
result$Gene <- NULL
result$sam <- NULL
sums <- colSums(result[,-1])
proportion_result <- sweep(result[,-1], 2, sums, '/' )
#proportion_result <- sapply(proportion_result, function(x) log(x))
proportion_result$Means <- rowMeans(proportion_result)
proportion_result$Target <- result$Target
y <- data.frame(Target = character(127))
y$Target <- result$Target
y$logresult <- log(proportion_result$IonXpress_8_Run_18_hg19_v3)
y$logMeans <- log(proportion_result$Means)
y$result <- proportion_result$IonXpress_8_Run_18_hg19_v3
y$Means <- proportion_result$Means
y[y$logMeans == -Inf,]$logMeans <- -10
y[y$logresult == -Inf,]$logresult <- -10
rownames(y) <- y$Target
```
Prepare features
```{r}
prepared_data <- data.frame(Target = character(172))
prepared_data$Target <- data$Amplicon_ID

prepared_data$Start_Dangling_End <- data$Start_Dangling_End
prepared_data$Stop_Dangling_End <- data$Stop_Dangling_End

prepared_data$Primer_Len <- as.numeric(lapply(as.character(data$Ion_AmpliSeq_Fwd_Primer.), function(x) nchar(x)))

prepared_data$Amplicon_Len <- as.numeric(lapply(as.character(data$Amplicon_Inner_Part), function(x) nchar(x)))
prepared_data$END_3 <- as.character(lapply(data$Amplicon_Inner_Part, function(x) substr(x, nchar(x), nchar(x))))
countCharOccurrences <- function(char, s) {s2 <- gsub(char,"",s);return (nchar(s) - nchar(s2))}
countGCcontent <- function(x){ gg <- countCharOccurrences('G',x); cc <-countCharOccurrences('C',x); aa <- countCharOccurrences('A',x); tt <- countCharOccurrences('T',x); (cc+gg)*100/(aa+tt+cc+gg) }

```


```{r}
prepared_data$Primer_GC_content <- as.numeric(lapply(as.character(data$Ion_AmpliSeq_Fwd_Primer.), function(x)  countGCcontent(x)))
prepared_data$Amplicon_GC_content <- as.numeric(lapply(as.character(data$Amplicon_Inner_Part), function(x) countGCcontent(x)))
rownames(prepared_data) <- prepared_data$Target
valid_names <- intersect(y$Target, data$Amplicon_ID)
prepared_data <- prepared_data[valid_names,]
prepared_data$Start_Dangling_End <- as.factor(prepared_data$Start_Dangling_End)
prepared_data$Stop_Dangling_End <- as.factor(prepared_data$Stop_Dangling_End)
prepared_data$END_3 <- as.factor(prepared_data$END_3)
y <- y[valid_names,]
```

```{r}
library(lattice)
bwplot(prepared_data$END_3 ~ y$logMeans,     ylab="nucleotids", xlab="log of amplicon percental amount", 
       main="3' end")
bwplot(prepared_data$Start_Dangling_End ~ y$logMeans,     ylab="nucleotids", xlab="log of amplicon percental amount", 
       main="First dangling end")
bwplot(prepared_data$Stop_Dangling_End ~ y$logMeans,     ylab="nucleotids", xlab="log of amplicon percental amount", 
       main="Second dangling end")
bwplot(prepared_data$END_3 ~ y$Means,     ylab="nucleotids", xlab="amplicon percental amount", 
        main="3' end")
bwplot(prepared_data$Start_Dangling_End ~ y$Means,     ylab="nucleotids", xlab="amplicon percental amount", 
      main="First dangling end")
bwplot(prepared_data$Stop_Dangling_End ~ y$Means,     ylab="nucleotids", xlab="amplicon percental amount", 
       main="Second dangling end")
xyplot(y$logMeans ~ prepared_data$Amplicon_GC_content,  xlab="GC-content(%)", ylab="log of amplicon percental amount", main="Amplicon GC-content")
xyplot(y$logMeans ~ prepared_data$Primer_GC_content,  xlab="GC-content(%)", ylab="log of amplicon percental amount", main="Primer GC-content")
xyplot(y$logMeans ~ prepared_data$Primer_Len,  xlab="Length", ylab="log of amplicon percental amount", main="Primer length")
xyplot(y$logMeans ~ prepared_data$Amplicon_Len,  xlab="Length", ylab="log of amplicon percental amount", main="Amplicon length")
xyplot(y$Means ~ prepared_data$Amplicon_GC_content,  xlab="GC-content(%)", ylab="amplicon percental amount", main="Amplicon GC-content")
xyplot(y$Means ~ prepared_data$Primer_GC_content,  xlab="GC-content(%)", ylab="amplicon percental amount", main="Amplicon GC-content")
xyplot(y$Means ~ prepared_data$Amplicon_Len,  xlab="Length", ylab="amplicon percental amount", main="Amplicon length")
xyplot(y$Means ~ prepared_data$Primer_Len,  xlab="Length", ylab="amplicon percental amount", main="Primer length")

histogram(~y$logMeans|factor(prepared_data$Start_Dangling_End))
```

```{r}
prepared_data$res <- y$Means
prepared_data$logres <- y$logMeans
write.table(prepared_data, "../data/features.tsv", col.names = TRUE, sep = "\t", row.names = FALSE)
```

Use all features to train model
```{r}
library(randomForest)
splitdf <- function(dataframe, n = 2, seed=NULL) {
    if (!is.null(seed)) set.seed(seed)
    index <- 1:nrow(dataframe)
    trainindex <- sample(index, trunc(length(index)*(n-1)/n))
    trainset <- dataframe[trainindex, ]
    testset <- dataframe[-trainindex, ]
    list(trainset=trainset,testset=testset)
}
splited <- splitdf(prepared_data)
trainSet <- splited$trainset
testSet <- splited$testset
formula <- res ~ Amplicon_Len + Amplicon_GC_content + END_3+Start_Dangling_End+Stop_Dangling_End + Primer_Len +Primer_GC_content
rf <- randomForest(formula, trainSet)
importance(rf)
fit <- lm(formula, data=trainSet)
#plot(fit)
predicted <- predict(fit, newdata = testSet)
rmse <- function(error)
{
    sqrt(mean(error^2))
}
mean(testSet$res)
sd(testSet$res)
rmse(predicted - testSet$res)
predicted <- predict(rf, newdata = testSet)
rmse(predicted - testSet$res)
```


Use only amplicon information and 3' end to train model
```{r}
formula <- res ~ Amplicon_Len + Amplicon_GC_content + END_3
rf <- randomForest(formula, trainSet)
importance(rf)
fit <- lm(formula, data=trainSet)
#plot(fit)
predicted <- predict(fit, newdata = testSet)
mean(testSet$res)
rmse(predicted - testSet$res)
predicted <- predict(rf, newdata = testSet)
rmse(predicted - testSet$res)
```



Use only amplicon information to train model
```{r}
formula <- res ~ Amplicon_Len + Amplicon_GC_content
rf <- randomForest(formula, trainSet)
importance(rf)
fit <- lm(formula, data=trainSet)
#plot(fit)
predicted <- predict(fit, newdata = testSet)
mean(testSet$res)
rmse(predicted - testSet$res)
predicted <- predict(rf, newdata = testSet)
rmse(predicted - testSet$res)
```


Use other to train model
```{r}
formula <- res ~ END_3+Start_Dangling_End+Stop_Dangling_End + Primer_Len +Primer_GC_content
rf <- randomForest(formula, trainSet)
importance(rf)
fit <- lm(formula, data=trainSet)
#plot(fit)
predicted <- predict(fit, newdata = testSet)
mean(testSet$res)
rmse(predicted - testSet$res)
predicted <- predict(rf, newdata = testSet)
rmse(predicted - testSet$res)
```
