---
title: "Amplicon similarity through experiments"
author: "tanunia"
date: "11/24/2014"
output: html_document
---

Load results and normalize results for every amplicon
```{r, echo=FALSE}
readResultFile <- function(name, addName = FALSE, delim = "\t"){
  filePath <- paste(c("../data/", name), collapse = "")
  result <- read.table(filePath, sep = delim, header = TRUE)
  result$Gene <- NULL
  result$sam <- NULL
  result$X <- NULL
  sums <- rowSums(result[,-1])
  proportion_result <- sweep(result[,-1], 1, sums, '/' )
  proportion_result$Target <- result$Target
  proportion_result
}
```

```{r}
y <- readResultFile("result.csv",FALSE, ",")
filePath <- paste(c("../data/", "result.csv"), collapse = "")
result <- read.table(filePath, sep = ",", header = TRUE)
result$Gene <- NULL
result$sam <- NULL
result$X <- NULL
sums <- rowSums(result[,-1])
y$sum <- sums
#resultFileNames <- c("SN1-17_Run_14.xls", "SN1-21_Run_17.xls","SN1-23-Run_18.xls","SN1-25-Run_19.xls", "SN1-26-Run_20.xls")
#resultTable <- data.frame(Target = character(127))
#for (i in 1:length(resultFileNames)){
#  resTable <- readResultFile(resultFileNames[i], TRUE)
#}
```

Simple KMeans clustering with a given number of clusters
```{r}
y <- y[complete.cases(y), ]
mydata <- y[, 1:(ncol(y)-2)]
fit <- kmeans(mydata, 4)
y$clusterKMeans <- fit$cluster
cat("KMeans given number of clusters: 4\n")
```

Choose number of clusters using Calinski-Harabasz criterion
```{r}
library(vegan)
fit <- cascadeKM(scale(mydata, center = TRUE,  scale = TRUE), 1, 8, iter = 1000)
plot(fit, sortg = TRUE, grpmts.plot = TRUE)
calinski.best <- as.numeric(which.max(fit$results[2,]))
cat("Calinski criterion optimal number of clusters:", calinski.best, "\n")
y$clusterCalinski <- fit$partition[,calinski.best]
```

Choose number of clusters using Affinity propagation
```{r}
library(apcluster)
d.apclus <- apcluster(negDistMat(r=2), mydata)
cat("Affinity propagation optimal number of clusters:", length(d.apclus@clusters), "\n")
y$clusterAP <- labels(d.apclus, type="enum")
```

Hierarchical clustering
```{r}
d_dist <- dist(as.matrix(mydata))
fit <- hclust(d_dist, method="ward") 
plot(fit)
groups <- cutree(fit, k=7)
rect.hclust(fit, k=7, border="red")
y$clusterHierarchy <- groups
```


Hierarchical clustering with correlation matrix
```{r}
t.mydata <- t(mydata)
colnames(t.mydata) <- y$Target
data.scale<- scale(t.mydata, center=TRUE, scale=TRUE)
cor_data <- cor(data.scale)
cor_d <- as.dist(cor_data)
fit <- hclust(cor_d, method="ward") 
plot(fit)
groups <- cutree(fit, k=7)
rect.hclust(fit, k=7, border="red")
y$clusterCor <- groups
```

Load prepared data
```{r}
prepared_data <- read.table("../data/features.tsv", sep = "\t", header=TRUE)
rownames(prepared_data) <- prepared_data$Target
rownames(y) <- y$Target
valid_names <- intersect(y$Target, prepared_data$Target)
prepared_data <- prepared_data[valid_names,]
y <- y[valid_names,]
```

----
### KMeans. Draw features for every cluster
```{r}
library(lattice)
data <- prepared_data
data$cluster <- as.factor(y$clusterKMeans)
histogram(~y$sum|data$cluster, main = "KMeans")
histogram(~data$Amplicon_GC_content|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Len|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Len|data$cluster, main = "KMeans")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Entropy|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Hash|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "KMeans")

histogram(~data$Fwd_END_3|data$cluster, main = "KMeans")
histogram(~data$Rev_END_3|data$cluster, main = "KMeans")

histogram(~data$Start_Dangling_End|data$cluster, main = "KMeans")
histogram(~data$Stop_Dangling_End|data$cluster, main = "KMeans")
```