---
title: "Amplicon similarity through experiments"
author: "tanunia"
date: "11/24/2014"
output: html_document
---

Load results and normalize results for every amplicon
```{r, echo=FALSE}
readResultFile <- function(name, addName = FALSE, delim = "\t"){
  filePath <- paste(c("../data/", name), collapse = "")
  result <- read.table(filePath, sep = delim, header = TRUE)
  result$Gene <- NULL
  result$sam <- NULL
  result$X <- NULL
  sums <- rowSums(result[,-1])
  proportion_result <- sweep(result[,-1], 1, sums, '/' )
  proportion_result$Target <- result$Target
  proportion_result
}
```

```{r}
y <- readResultFile("result.csv",FALSE, ",")
filePath <- paste(c("../data/", "result.csv"), collapse = "")
result <- read.table(filePath, sep = ",", header = TRUE)
result$Gene <- NULL
result$sam <- NULL
result$X <- NULL
sums <- rowSums(result[,-1])
y$sum <- sums
#resultFileNames <- c("SN1-17_Run_14.xls", "SN1-21_Run_17.xls","SN1-23-Run_18.xls","SN1-25-Run_19.xls", "SN1-26-Run_20.xls")
#resultTable <- data.frame(Target = character(127))
#for (i in 1:length(resultFileNames)){
#  resTable <- readResultFile(resultFileNames[i], TRUE)
#}
```

Simple KMeans clustering with a given number of clusters
```{r}
y <- y[complete.cases(y), ]
mydata <- y[, 1:(ncol(y)-2)]
fit <- kmeans(mydata, 4)
y$clusterKMeans <- fit$cluster
cat("KMeans given number of clusters: 4\n")
```

Choose number of clusters using Calinski-Harabasz criterion
```{r}
library(vegan)
fit <- cascadeKM(scale(mydata, center = TRUE,  scale = TRUE), 1, 8, iter = 1000)
plot(fit, sortg = TRUE, grpmts.plot = TRUE)
calinski.best <- as.numeric(which.max(fit$results[2,]))
cat("Calinski criterion optimal number of clusters:", calinski.best, "\n")
y$clusterCalinski <- fit$partition[,calinski.best]
```

Choose number of clusters using Affinity propagation
```{r}
library(apcluster)
d.apclus <- apcluster(negDistMat(r=2), mydata)
cat("Affinity propagation optimal number of clusters:", length(d.apclus@clusters), "\n")
y$clusterAP <- labels(d.apclus, type="enum")
```

Hierarchical clustering
```{r}
d_dist <- dist(as.matrix(mydata))
fit <- hclust(d_dist, method="ward") 
plot(fit)
groups <- cutree(fit, k=7)
rect.hclust(fit, k=7, border="red")
y$clusterHierarchy <- groups
```


Hierarchical clustering with correlation matrix
```{r}
t.mydata <- t(mydata)
colnames(t.mydata) <- y$Target
data.scale<- scale(t.mydata, center=TRUE, scale=TRUE)
cor_data <- cor(data.scale)
cor_d <- as.dist(cor_data)
fit <- hclust(cor_d, method="ward") 
plot(fit)
groups <- cutree(fit, k=7)
rect.hclust(fit, k=7, border="red")
y$clusterCor <- groups
```

Load prepared data
```{r}
prepared_data <- read.table("../data/features.tsv", sep = "\t", header=TRUE)
rownames(prepared_data) <- prepared_data$Target
rownames(y) <- y$Target
valid_names <- intersect(y$Target, prepared_data$Target)
prepared_data <- prepared_data[valid_names,]
y <- y[valid_names,]
```

----
### KMeans. Draw features for every cluster
```{r}
library(lattice)
data <- prepared_data
data$cluster <- as.factor(y$clusterKMeans)
#histogram(~y$sum|data$cluster, main = "KMeans")
histogram(~data$Amplicon_GC_content|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Len|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Len|data$cluster, main = "KMeans")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Entropy|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "KMeans")

histogram(~data$Amplicon_Hash|data$cluster, main = "KMeans")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "KMeans")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "KMeans")

histogram(~data$Fwd_END_3|data$cluster, main = "KMeans")
histogram(~data$Rev_END_3|data$cluster, main = "KMeans")

histogram(~data$Start_Dangling_End|data$cluster, main = "KMeans")
histogram(~data$Stop_Dangling_End|data$cluster, main = "KMeans")
```
----
### Calinski criterion. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterCalinski)
histogram(~data$Amplicon_GC_content|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Amplicon_Len|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_Primer_Len|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Amplicon_Entropy|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Amplicon_Hash|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Fwd_END_3|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Rev_END_3|data$cluster, main = "KMeans.Calinski-Harabasz criterion")

histogram(~data$Start_Dangling_End|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
histogram(~data$Stop_Dangling_End|data$cluster, main = "KMeans.Calinski-Harabasz criterion")
```
----
### Affinity propagation. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterAP)
histogram(~data$Amplicon_GC_content|data$cluster, main = "Affinity propagation")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "Affinity propagation")

histogram(~data$Amplicon_Len|data$cluster, main = "Affinity propagation")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_Primer_Len|data$cluster, main = "Affinity propagation")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "Affinity propagation")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "Affinity propagation")

histogram(~data$Amplicon_Entropy|data$cluster, main = "Affinity propagation")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "Affinity propagation")

histogram(~data$Amplicon_Hash|data$cluster, main = "Affinity propagation")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "Affinity propagation")

histogram(~data$Fwd_END_3|data$cluster, main = "Affinity propagation")
histogram(~data$Rev_END_3|data$cluster, main = "Affinity propagation")

histogram(~data$Start_Dangling_End|data$cluster, main = "Affinity propagation")
histogram(~data$Stop_Dangling_End|data$cluster, main = "Affinity propagation")
```
----
### Hierarchical clustering. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterHierarchy)
histogram(~data$Amplicon_GC_content|data$cluster, main = "Hierarchical clustering")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "Hierarchical clustering")

histogram(~data$Amplicon_Len|data$cluster, main = "Hierarchical clustering")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_Primer_Len|data$cluster, main = "Hierarchical clustering")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "Hierarchical clustering")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "Hierarchical clustering")

histogram(~data$Amplicon_Entropy|data$cluster, main = "Hierarchical clustering")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "Hierarchical clustering")

histogram(~data$Amplicon_Hash|data$cluster, main = "Hierarchical clustering")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "Hierarchical clustering")

histogram(~data$Fwd_END_3|data$cluster, main = "Hierarchical clustering")
histogram(~data$Rev_END_3|data$cluster, main = "Hierarchical clustering")

histogram(~data$Start_Dangling_End|data$cluster, main = "Hierarchical clustering")
histogram(~data$Stop_Dangling_End|data$cluster, main = "Hierarchical clustering")
```
----
### Correlation clustering. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterCor)
histogram(~data$Amplicon_GC_content|data$cluster, main = "Correlation clustering")
histogram(~data$Fwd_Primer_GC_content|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_Primer_GC_content|data$cluster, main = "Correlation clustering")

histogram(~data$Amplicon_Len|data$cluster, main = "Correlation clustering")
histogram(~data$Fwd_Primer_Len|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_Primer_Len|data$cluster, main = "Correlation clustering")

histogram(~data$Amplicon_MaxPolyX|data$cluster, main = "Correlation clustering")
histogram(~data$Fwd_Primer_MaxPolyX|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_Primer_MaxPolyX|data$cluster, main = "Correlation clustering")

histogram(~data$Amplicon_Entropy|data$cluster, main = "Correlation clustering")
histogram(~data$Fwd_Primer_Entropy|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_Primer_Entropy|data$cluster, main = "Correlation clustering")

histogram(~data$Amplicon_Hash|data$cluster, main = "Correlation clustering")
histogram(~data$Fwd_Primer_Hash|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_Primer_Hash|data$cluster, main = "Correlation clustering")

histogram(~data$Fwd_END_3|data$cluster, main = "Correlation clustering")
histogram(~data$Rev_END_3|data$cluster, main = "Correlation clustering")

histogram(~data$Start_Dangling_End|data$cluster, main = "Correlation clustering")
histogram(~data$Stop_Dangling_End|data$cluster, main = "Correlation clustering")
```

-----
-----
### KMeans. Draw features for every cluster
```{r, echo=FALSE}
getFactoredDataTrunc <- function(dt){
    q <- quantile(dt, probs = c(1/4, 1/2, 3/4, 1))
    print(q)
    dt[dt <= q[1]] <- trunc(q[1])
    dt[dt > q[1] & dt <= q[2]] <- trunc(q[2])
    dt[dt > q[2] & dt <= q[3]] <- trunc(q[3])
    dt[dt > q[3]] <- trunc(q[4])
    dt
}
```
```{r, echo=FALSE}
getFactoredData <- function(dt){
    q <- quantile(dt, probs = c(1/4, 1/2, 3/4, 1))
    print(q)
    dt[dt <= q[1]] <- q[1]
    dt[dt > q[1] & dt <= q[2]] <- q[2]
    dt[dt > q[2] & dt <= q[3]] <- q[3]
    dt[dt > q[3]] <- q[4]
    dt
}
```
```{r}
library(lattice)
data <- prepared_data
data$cluster <- as.factor(y$clusterKMeans)
histogram(data$cluster)
amplicon_gc_content <- getFactoredDataTrunc(data$Amplicon_GC_content)
counts <- table(data$cluster, amplicon_gc_content)
barplot(counts, main="Amplicon Distribution by Amplicon GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_gc_content <- getFactoredDataTrunc(data$Fwd_Primer_GC_content)
counts <- table(data$cluster, fwd_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Fwd Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_gc_content <- getFactoredDataTrunc(data$Rev_Primer_GC_content)
counts <- table(data$cluster, rev_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Rev Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_len <- getFactoredDataTrunc(data$Amplicon_Len)
counts <- table(data$cluster, amplicon_len)
barplot(counts, main="Amplicon Distribution by Amplicon Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_len <- getFactoredDataTrunc(data$Fwd_Primer_Len)
counts <- table(data$cluster, fwd_primer_len)
barplot(counts, main="Amplicon Distribution by Fwd Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_len <- getFactoredDataTrunc(data$Rev_Primer_Len)
counts <- table(data$cluster, rev_primer_len)
barplot(counts, main="Amplicon Distribution by Rev Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_MaxPolyX <- getFactoredDataTrunc(data$Amplicon_MaxPolyX)
counts <- table(data$cluster, amplicon_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Amplicon MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_MaxPolyX <- getFactoredDataTrunc(data$Fwd_Primer_MaxPolyX)
counts <- table(data$cluster, fwd_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Fwd Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_MaxPolyX <- getFactoredDataTrunc(data$Rev_Primer_MaxPolyX)
counts <- table(data$cluster, rev_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Rev Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_Entropy <- getFactoredData(data$Amplicon_Entropy)
counts <- table(data$cluster, amplicon_Entropy)
barplot(counts, main="Amplicon Distribution by Amplicon Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Entropy <- getFactoredData(data$Fwd_Primer_Entropy)
counts <- table(data$cluster, fwd_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Fwd Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Entropy <- getFactoredData(data$Rev_Primer_Entropy)
counts <- table(data$cluster, rev_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Rev Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_Hash <- getFactoredDataTrunc(data$Amplicon_Hash)
counts <- table(data$cluster, amplicon_Hash)
barplot(counts, main="Amplicon Distribution by Amplicon Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Hash <- getFactoredDataTrunc(data$Fwd_Primer_Hash)
counts <- table(data$cluster, fwd_primer_Hash)
barplot(counts, main="Amplicon Distribution by Fwd Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Hash <- getFactoredDataTrunc(data$Rev_Primer_Hash)
counts <- table(data$cluster, rev_primer_Hash)
barplot(counts, main="Amplicon Distribution by Rev Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

histogram(~data$cluster|data$Fwd_END_3, main = "KMeans")
histogram(~data$cluster|data$Rev_END_3, main = "KMeans")

histogram(~data$cluster|data$Start_Dangling_End, main = "KMeans")
histogram(~data$cluster|data$Stop_Dangling_End, main = "KMeans")
```
----
### Calinski criterion. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterCalinski)
histogram(data$cluster)
amplicon_gc_content <- getFactoredDataTrunc(data$Amplicon_GC_content)
counts <- table(data$cluster, amplicon_gc_content)
barplot(counts, main="Amplicon Distribution by Amplicon GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_gc_content <- getFactoredDataTrunc(data$Fwd_Primer_GC_content)
counts <- table(data$cluster, fwd_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Fwd Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_gc_content <- getFactoredDataTrunc(data$Rev_Primer_GC_content)
counts <- table(data$cluster, rev_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Rev Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_len <- getFactoredDataTrunc(data$Amplicon_Len)
counts <- table(data$cluster, amplicon_len)
barplot(counts, main="Amplicon Distribution by Amplicon Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_len <- getFactoredDataTrunc(data$Fwd_Primer_Len)
counts <- table(data$cluster, fwd_primer_len)
barplot(counts, main="Amplicon Distribution by Fwd Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_len <- getFactoredDataTrunc(data$Rev_Primer_Len)
counts <- table(data$cluster, rev_primer_len)
barplot(counts, main="Amplicon Distribution by Rev Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_MaxPolyX <- getFactoredDataTrunc(data$Amplicon_MaxPolyX)
counts <- table(data$cluster, amplicon_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Amplicon MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_MaxPolyX <- getFactoredDataTrunc(data$Fwd_Primer_MaxPolyX)
counts <- table(data$cluster, fwd_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Fwd Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_MaxPolyX <- getFactoredDataTrunc(data$Rev_Primer_MaxPolyX)
counts <- table(data$cluster, rev_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Rev Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_Entropy <- getFactoredData(data$Amplicon_Entropy)
counts <- table(data$cluster, amplicon_Entropy)
barplot(counts, main="Amplicon Distribution by Amplicon Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Entropy <- getFactoredData(data$Fwd_Primer_Entropy)
counts <- table(data$cluster, fwd_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Fwd Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Entropy <- getFactoredData(data$Rev_Primer_Entropy)
counts <- table(data$cluster, rev_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Rev Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_Hash <- getFactoredDataTrunc(data$Amplicon_Hash)
counts <- table(data$cluster, amplicon_Hash)
barplot(counts, main="Amplicon Distribution by Amplicon Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Hash <- getFactoredDataTrunc(data$Fwd_Primer_Hash)
counts <- table(data$cluster, fwd_primer_Hash)
barplot(counts, main="Amplicon Distribution by Fwd Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Hash <- getFactoredDataTrunc(data$Rev_Primer_Hash)
counts <- table(data$cluster, rev_primer_Hash)
barplot(counts, main="Amplicon Distribution by Rev Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


histogram(~data$cluster|data$Fwd_END_3, main =  "KMeans.Calinski-Harabasz criterion")
histogram(~data$cluster|data$Rev_END_3, main =  "KMeans.Calinski-Harabasz criterion")

histogram(~data$cluster|data$Start_Dangling_End, main =  "KMeans.Calinski-Harabasz criterion")
histogram(~data$cluster|data$Stop_Dangling_End, main =  "KMeans.Calinski-Harabasz criterion")
```
----
### Affinity propagation. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterAP)
histogram(data$cluster)
amplicon_gc_content <- getFactoredDataTrunc(data$Amplicon_GC_content)
counts <- table(data$cluster, amplicon_gc_content)
barplot(counts, main="Amplicon Distribution by Amplicon GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_gc_content <- getFactoredDataTrunc(data$Fwd_Primer_GC_content)
counts <- table(data$cluster, fwd_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Fwd Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_gc_content <- getFactoredDataTrunc(data$Rev_Primer_GC_content)
counts <- table(data$cluster, rev_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Rev Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_len <- getFactoredDataTrunc(data$Amplicon_Len)
counts <- table(data$cluster, amplicon_len)
barplot(counts, main="Amplicon Distribution by Amplicon Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_len <- getFactoredDataTrunc(data$Fwd_Primer_Len)
counts <- table(data$cluster, fwd_primer_len)
barplot(counts, main="Amplicon Distribution by Fwd Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_len <- getFactoredDataTrunc(data$Rev_Primer_Len)
counts <- table(data$cluster, rev_primer_len)
barplot(counts, main="Amplicon Distribution by Rev Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_MaxPolyX <- getFactoredDataTrunc(data$Amplicon_MaxPolyX)
counts <- table(data$cluster, amplicon_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Amplicon MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_MaxPolyX <- getFactoredDataTrunc(data$Fwd_Primer_MaxPolyX)
counts <- table(data$cluster, fwd_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Fwd Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_MaxPolyX <- getFactoredDataTrunc(data$Rev_Primer_MaxPolyX)
counts <- table(data$cluster, rev_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Rev Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_Entropy <- getFactoredData(data$Amplicon_Entropy)
counts <- table(data$cluster, amplicon_Entropy)
barplot(counts, main="Amplicon Distribution by Amplicon Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Entropy <- getFactoredData(data$Fwd_Primer_Entropy)
counts <- table(data$cluster, fwd_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Fwd Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Entropy <- getFactoredData(data$Rev_Primer_Entropy)
counts <- table(data$cluster, rev_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Rev Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_Hash <- getFactoredDataTrunc(data$Amplicon_Hash)
counts <- table(data$cluster, amplicon_Hash)
barplot(counts, main="Amplicon Distribution by Amplicon Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Hash <- getFactoredDataTrunc(data$Fwd_Primer_Hash)
counts <- table(data$cluster, fwd_primer_Hash)
barplot(counts, main="Amplicon Distribution by Fwd Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Hash <- getFactoredDataTrunc(data$Rev_Primer_Hash)
counts <- table(data$cluster, rev_primer_Hash)
barplot(counts, main="Amplicon Distribution by Rev Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


histogram(~data$cluster|data$Fwd_END_3, main = "Affinity propagation")
histogram(~data$cluster|data$Rev_END_3, main = "Affinity propagation")

histogram(~data$cluster|data$Start_Dangling_End, main = "Affinity propagation")
histogram(~data$cluster|data$Stop_Dangling_End, main = "Affinity propagation")
```
----
### Hierarchical clustering. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterHierarchy)
histogram(data$cluster)
amplicon_gc_content <- getFactoredDataTrunc(data$Amplicon_GC_content)
counts <- table(data$cluster, amplicon_gc_content)
barplot(counts, main="Amplicon Distribution by Amplicon GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_gc_content <- getFactoredDataTrunc(data$Fwd_Primer_GC_content)
counts <- table(data$cluster, fwd_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Fwd Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_gc_content <- getFactoredDataTrunc(data$Rev_Primer_GC_content)
counts <- table(data$cluster, rev_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Rev Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_len <- getFactoredDataTrunc(data$Amplicon_Len)
counts <- table(data$cluster, amplicon_len)
barplot(counts, main="Amplicon Distribution by Amplicon Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_len <- getFactoredDataTrunc(data$Fwd_Primer_Len)
counts <- table(data$cluster, fwd_primer_len)
barplot(counts, main="Amplicon Distribution by Fwd Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_len <- getFactoredDataTrunc(data$Rev_Primer_Len)
counts <- table(data$cluster, rev_primer_len)
barplot(counts, main="Amplicon Distribution by Rev Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_MaxPolyX <- getFactoredDataTrunc(data$Amplicon_MaxPolyX)
counts <- table(data$cluster, amplicon_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Amplicon MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_MaxPolyX <- getFactoredDataTrunc(data$Fwd_Primer_MaxPolyX)
counts <- table(data$cluster, fwd_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Fwd Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_MaxPolyX <- getFactoredDataTrunc(data$Rev_Primer_MaxPolyX)
counts <- table(data$cluster, rev_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Rev Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_Entropy <- getFactoredData(data$Amplicon_Entropy)
counts <- table(data$cluster, amplicon_Entropy)
barplot(counts, main="Amplicon Distribution by Amplicon Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Entropy <- getFactoredData(data$Fwd_Primer_Entropy)
counts <- table(data$cluster, fwd_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Fwd Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Entropy <- getFactoredData(data$Rev_Primer_Entropy)
counts <- table(data$cluster, rev_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Rev Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_Hash <- getFactoredDataTrunc(data$Amplicon_Hash)
counts <- table(data$cluster, amplicon_Hash)
barplot(counts, main="Amplicon Distribution by Amplicon Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Hash <- getFactoredDataTrunc(data$Fwd_Primer_Hash)
counts <- table(data$cluster, fwd_primer_Hash)
barplot(counts, main="Amplicon Distribution by Fwd Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Hash <- getFactoredDataTrunc(data$Rev_Primer_Hash)
counts <- table(data$cluster, rev_primer_Hash)
barplot(counts, main="Amplicon Distribution by Rev Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


histogram(~data$cluster|data$Fwd_END_3, main = "Hierarchical clustering")
histogram(~data$cluster|data$Rev_END_3, main = "Hierarchical clustering")

histogram(~data$cluster|data$Start_Dangling_End, main = "Hierarchical clustering")
histogram(~data$cluster|data$Stop_Dangling_End, main = "Hierarchical clustering")
```
----
### Correlation clustering. Draw features for every cluster
```{r}
data$cluster <- as.factor(y$clusterCor)
histogram(data$cluster)
amplicon_gc_content <- getFactoredDataTrunc(data$Amplicon_GC_content)
counts <- table(data$cluster, amplicon_gc_content)
barplot(counts, main="Amplicon Distribution by Amplicon GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_gc_content <- getFactoredDataTrunc(data$Fwd_Primer_GC_content)
counts <- table(data$cluster, fwd_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Fwd Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_gc_content <- getFactoredDataTrunc(data$Rev_Primer_GC_content)
counts <- table(data$cluster, rev_primer_gc_content)
barplot(counts, main="Amplicon Distribution by Rev Primer GC-content and Clusters",
  xlab="GC-content", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_len <- getFactoredDataTrunc(data$Amplicon_Len)
counts <- table(data$cluster, amplicon_len)
barplot(counts, main="Amplicon Distribution by Amplicon Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_len <- getFactoredDataTrunc(data$Fwd_Primer_Len)
counts <- table(data$cluster, fwd_primer_len)
barplot(counts, main="Amplicon Distribution by Fwd Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_len <- getFactoredDataTrunc(data$Rev_Primer_Len)
counts <- table(data$cluster, rev_primer_len)
barplot(counts, main="Amplicon Distribution by Rev Primer Length and Clusters",
  xlab="Length", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_MaxPolyX <- getFactoredDataTrunc(data$Amplicon_MaxPolyX)
counts <- table(data$cluster, amplicon_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Amplicon MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_MaxPolyX <- getFactoredDataTrunc(data$Fwd_Primer_MaxPolyX)
counts <- table(data$cluster, fwd_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Fwd Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_MaxPolyX <- getFactoredDataTrunc(data$Rev_Primer_MaxPolyX)
counts <- table(data$cluster, rev_primer_MaxPolyX)
barplot(counts, main="Amplicon Distribution by Rev Primer MaxPolyX and Clusters",
  xlab="MaxPolyX", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


amplicon_Entropy <- getFactoredData(data$Amplicon_Entropy)
counts <- table(data$cluster, amplicon_Entropy)
barplot(counts, main="Amplicon Distribution by Amplicon Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Entropy <- getFactoredData(data$Fwd_Primer_Entropy)
counts <- table(data$cluster, fwd_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Fwd Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Entropy <- getFactoredData(data$Rev_Primer_Entropy)
counts <- table(data$cluster, rev_primer_Entropy)
barplot(counts, main="Amplicon Distribution by Rev Primer Entropy and Clusters",
  xlab="Entropy", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))

amplicon_Hash <- getFactoredDataTrunc(data$Amplicon_Hash)
counts <- table(data$cluster, amplicon_Hash)
barplot(counts, main="Amplicon Distribution by Amplicon Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
fwd_primer_Hash <- getFactoredDataTrunc(data$Fwd_Primer_Hash)
counts <- table(data$cluster, fwd_primer_Hash)
barplot(counts, main="Amplicon Distribution by Fwd Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))
rev_primer_Hash <- getFactoredDataTrunc(data$Rev_Primer_Hash)
counts <- table(data$cluster, rev_primer_Hash)
barplot(counts, main="Amplicon Distribution by Rev Primer Hash and Clusters",
  xlab="Hash", col=c("darkblue","red","green","blue"),
   legend = rownames(counts))


histogram(~data$cluster|data$Fwd_END_3, main = "Correlation clustering")
histogram(~data$cluster|data$Rev_END_3, main = "Correlation clustering")

histogram(~data$cluster|data$Start_Dangling_End, main = "Correlation clustering")
histogram(~data$cluster|data$Stop_Dangling_End, main = "Correlation clustering")
```


### Amplification results
```{r, echo=FALSE}
getFactoredDataTrunc <- function(dt){
    q <- quantile(dt, probs = c(1/4, 1/2, 3/4, 1))
    print(q)
    dt[dt <= q[1]] <- trunc(q[1])
    dt[dt > q[1] & dt <= q[2]] <- trunc(q[2])
    dt[dt > q[2] & dt <= q[3]] <- trunc(q[3])
    dt[dt > q[3]] <- trunc(q[4])
    dt
}
```

